The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-05-08 Paper/jaccardbaetica.tex
add changes and add image draft2

\address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}
\address[ceipacadress]{CEIPAC, Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}

@@

@@

Stamps are the most studied in this type of amphorae. The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning. Its endure in time become an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016} (Fig.\ref{amphora}).

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/workshop1}
\caption{Distribution of the workshops in Baetica province}
\label{workshop}
\end{figure}

The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
millet_anforas_1998}. This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}. Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown. Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.

\xavi{Estás segura de esto? Porque aquí hay 2 cosas: 1. datar bien y 2 . período de uso. Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo. En qué está basada la idea? Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}
Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology. This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.

@@

The frequency of stamps per workshop can be seen in Fig. \ref{stamps}. Most stamps are distributed in only one workshop whereas few workshops concentrate a high frequency of different amphoric stamps. The type of distribution is also frequent in amphora production where we observed a self-organized complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.

@@

\xavi{que me quede claro...la mayoría de workshops solo tienen 1 sello??? Y hay un workshop con como 700 sellos distintos??? }

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
\caption{Histogram with the number of stamps (X axis) and the frequency of workshop (Y axis). The distribution is widely diverse with most workshops have only one stamp}
\label{stamps}
\end{figure}

\xavi{si el eje es logarítmico deberías mencionarlo}

The distribution of amphora stamps in different \textit{conventus} can be seen in Fig. \ref{frequency}. The majority of stamps found are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively. Mostly, workshops show a homogeneity on the frequency of stamps except for both La Catria and Arva that exhibit a big amount of stamps with 29 different stamps. According to previous studies, those workshops became in most important centres of amphora production although they could have more intensely prospected than others. \citep{arva_1997}.

>>>>

\address[ceipacadress]{CEIPAC, Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}
\address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/workshop1}
\caption{Distribution of the workshops (dots) in Baetica province}
\label{workshop}
\end{figure}

The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
millet_anforas_1998}. In particular, the importance of olive oil production could be explained by the fact that we detected amphorae distributed in different Roman provinces such as \textit{Germania}, \textit{Britannia} and \textit{Gallia}. This amphora was commonly associated with the  transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.

Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown. Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.

%\xavi{Estás segura de esto? Porque aquí hay 2 cosas: 1. datar bien y 2 . período de uso. Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo. En qué está basada la idea? Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}

%Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology in some cases. This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.

@@

Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}. The  period of use is associated with a large amphora production for almost three centuries.

Most stamps display a large activity of production that it can be difficult for specifying an accurate chronology in some cases. They have been frequently dated with the consultar year by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
However, the chronology of stamps by using \textit{tituli picti} could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.  
\maria{no sé cómo explicarlo mejor}

The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning. Its endure in time become an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016} (Fig.\ref{amphora}).

@@

@@

The frequency of stamps per workshop can be seen in Fig. \ref{stamps}. Most 
workshop contained only one stamp whereas one workshop (La Catria) concentrated a high frequency of amphoric stamps with a total number of 228 stamps. The type of distribution is also frequent in amphora production where we observed a self-organized complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.

@@

%\xavi{que me quede claro...la mayoría de workshops solo tienen 1 sello??? Y hay un workshop con como 700 sellos distintos??? }

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
\caption{Histogram on a log scale with base 10. X axis is represented by the number of stamps and Y axis is the frequency of workshop. The distribution is widely diverse with most workshops have only one stamp}
\label{stamps}
\end{figure}

%\xavi{si el eje es logarítmico deberías mencionarlo}

The distribution of amphora stamps in different \textit{conventus} can be seen in Fig. \ref{frequency}. The majority of stamps found are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively. Mostly, workshops show a homogeneity on the frequency of stamps except for both La Catria and Arva that exhibit a big amount of stamps with 29 different code stamps. According to previous studies, those workshops became in most important centres of amphora production although they could have more intensely prospected than others. \citep{arva_1997}.
